---
title: "Lab Examples, code and content"
output: 
  bookdown::github_document2
always_allow_html: true
---

The repository provides some examples of how to create your lab write-ups such that I can view and provide feedback within GitHub Classroom.

The first few lines of your `.Rmd` file will be a YAML. They should look something like this:

```{html}
---
title: "Title"
author: "Your name"
date: "the date"
output: github_document
---
```

There are many other options you can add to your YAML, but these are the basics. Some others that I use quite often are `toc: "true"` to add a table of contents and `urlcolor: "cyan"` to differentiate hyperlinks by coloring them light blue.

You can now complete your writeup in between "chunks" of code (see the `README.Rmd`).

# Examples

Below are a bunch of examples of things you can add to your report. The code for creating them can be found in `README.md`.

## Plots


```{r plotname, fig.height=3, fig.width=3, fig.cap="Example plot."}
library(tidyverse)

# create a plot
  ggplot(data = pressure, aes(x = temperature, y = pressure)) +
    geom_point() +
    theme_classic() +
    geom_smooth(se = F)
```

You can reference a plot, such as Figure \@ref(fig:plotname), using the syntax `\@ref(fig:plotname)` inside the markdown text if you install the `bookdown` R package and include the following in the YAML:

```{html}
---
output:
  bookdown::github_document2
---
```

## Tables

Tables can be rendered straight from dataframes using the `kableExtra` package.

```{r}
library(kableExtra)
# pass a few lines of a dataframe to kable
  head(pressure) %>%
    kable(
      caption = "An example table.",
      label = "eg-tab"
    ) %>%
    kable_classic(full_width = FALSE) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover"), full_width = FALSE
    )

```

Tables such as Table \@ref(tab:eg-tab) can also be referred to in the text by referencing the label supplied to the `label` argument using the syntax `\@ref(tab:label)`. 

**Note**: You may need to add `always_allow_html: true` to the YAML to get some tables generated using `kableExtra` to render in a `github_document`.

## Hiding some R code

You can always supply the option `echo=FALSE` to a code chunk if you do not want it to render. Keep your writeups short and to the point. Don't include any extaneous information or R code!

## Other

### Images

`![](path-to-image)`

### Hyperlinks

`[text to disply](url)`

### Math

You can add math symbols using LaTeX syntax. For example,

```{}
$$
  e^{i\pi} = -1
$$
```

will render as a display equation like so:

$$
e^{i\pi} = -1
$$
while

```{}
$e^{ix} = \cos(x) + i \sin(x)$
```

will render inline like so: $e^{ix} = \cos(x) + i \sin(x)$
